% ============================================================
% COMPARATIVE RESULTS -- IEEE Trans. Smart Grid (Journal Extension)
% Energy-Information Co-Optimization via Learnable Coupling
% ============================================================
%
% PURPOSE
% -------
% This file is a standalone compilable reference document that pulls in
% the auto-generated experiment tables (tables/table01 .. table10).
% It does NOT hard-code any numerical results; all numbers come from
% the \input-ed table files so they stay in sync with experiment reruns.
%
% After the K-scaling fix (k_init = safety_factor * |lambda_min| / n_gen),
% the old SmartGridComm 2026 numbers are obsolete.  This document
% supersedes the previous comparative_results.tex.
%
% MODELS (8 total)
% ----------------
% JointOptimizer (Ours), B1: Sequential OPF+QoS, B2: MLP Joint,
% B3: GNN Only, B4: LSTM Joint, B5: CNN Joint,
% B6: Vanilla Transformer, B7: Transformer No Coupling.
% Additional baselines in specific tables: B8: HeteroGNN, B9: DeepOPF.
%
% IEEE TEST CASES (5 + synthetic)
% -------------------------------
% IEEE 14-bus, 30-bus, 39-bus, 57-bus, 118-bus.
% Synthetic 10 000-bus used in selected experiments (inference, convergence).
%
% TRAINING ENVIRONMENT
% --------------------
% All experiments trained on GPU (CUDA).  K_init is auto-scaled per grid:
%   k_init = safety_factor * |lambda_min| / n_gen
%
% COMPILE
% -------
%   pdflatex comparative_results.tex
% ============================================================

\documentclass[journal]{IEEEtran}

\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{multirow}

\begin{document}

\title{Comparative Results Reference\\[4pt]
\large IEEE Trans. Smart Grid -- Journal Extension}
\author{Auto-generated tables from experiment pipeline}
\maketitle

% ============================================================
\section*{How to use this document}
% ============================================================
Each section below includes one auto-generated table via
\verb|\input{tables/tableXX_xxx.tex}| together with brief commentary
explaining what the table measures, which experiments produced it,
and what to look for when interpreting the numbers.

All tables were generated after the K-scaling fix
($K_{\mathrm{init}} = \mathit{safety\_factor} \cdot |\lambda_{\min}| / n_{\mathrm{gen}}$),
and reflect 5 IEEE test cases (14, 30, 39, 57, 118-bus) with 8 models
(JointOptimizer + B1--B7).  Training was performed on GPU (CUDA),
not on CPU as in the earlier conference submission.


% ============================================================
\section{Table I -- Main Baseline Comparison}
\label{sec:table01}
% ============================================================

\input{tables/table01_main_comparison.tex}

\medskip\noindent\textbf{What it shows.}
Stability rate (\%) and mean stability margin $\rho(\tau)$ (with standard
deviation over 5 random seeds) for all 8 models across all 5 IEEE cases.
This is the primary result table of the paper.

\noindent\textbf{Key points.}
\begin{itemize}
  \item JointOptimizer achieves the highest $\rho(\tau)$ on every case.
  \item All models achieve 100\% stability rate, so the differentiator is
        the \emph{margin} above the stability threshold.
  \item The improvement is most pronounced on larger grids where the
        auto-scaled $K_{\mathrm{init}}$ matters most.
\end{itemize}


% ============================================================
\section{Table II -- Ablation Study}
\label{sec:table02}
% ============================================================

\input{tables/table02_ablation.tex}

\medskip\noindent\textbf{What it shows.}
Impact of individual architectural components on the stability margin,
averaged across IEEE 14--118.  Ablated factors: coupling weight~$\alpha$,
physics mask, causal mask, cross-attention, contrastive loss
(Observation~1, formerly Theorem~2), and GNN depth.

\noindent\textbf{Key points.}
\begin{itemize}
  \item The coupling weight $\alpha$ is the most influential hyperparameter;
        disabling it ($\alpha=0$) recovers B7-level performance.
  \item Physics mask, causal mask, and cross-attention show marginal
        individual impact but contribute jointly (see contrastive ablation).
  \item GNN depth shows diminishing returns beyond 2--3 layers.
\end{itemize}


% ============================================================
\section{Table III -- Stress Test}
\label{sec:table03}
% ============================================================

\input{tables/table03_stress_test.tex}

\medskip\noindent\textbf{What it shows.}
Stability rate under progressively stressed operating conditions: increased
load (110\%, 120\% of nominal), combined moderate stress, and combined
severe stress.

\noindent\textbf{Key points.}
\begin{itemize}
  \item Evaluates robustness to out-of-distribution operating conditions.
  \item Stressed scenarios combine increased load with degraded
        communication quality.
  \item Stability degradation under severe stress reveals which
        architectures are most brittle.
\end{itemize}


% ============================================================
\section{Table IV -- Transfer Learning}
\label{sec:table04}
% ============================================================

\input{tables/table04_transfer.tex}

\medskip\noindent\textbf{What it shows.}
Cross-case generalization: a model trained on one IEEE case is evaluated
zero-shot and after fine-tuning on a different case.  Reports
$\rho(\tau)$ for both, stability rate, and relative improvement from
fine-tuning.

\noindent\textbf{Key points.}
\begin{itemize}
  \item Demonstrates that learned coupling structure transfers across
        grid topologies.
  \item Fine-tuning yields the largest improvement when source and target
        cases differ substantially in size (e.g., 39 $\to$ 118).
  \item Zero-shot transfer already preserves stability, confirming that
        the physics-informed features generalize.
\end{itemize}


% ============================================================
\section{Table V -- Inference Latency and Model Parameters}
\label{sec:table05}
% ============================================================

\input{tables/table05_inference.tex}

\medskip\noindent\textbf{What it shows.}
Inference latency (mean, P95, P99 in milliseconds) and parameter count
for JointOptimizer and selected baselines (B2, B3, B6, B8: HeteroGNN,
B9: DeepOPF) across all IEEE cases including the 10\,000-bus synthetic grid.

\noindent\textbf{Key points.}
\begin{itemize}
  \item JointOptimizer provides sub-50\,ms latency on cases up to
        10\,000 buses, well within real-time SCADA requirements.
  \item B8 (HeteroGNN) is orders of magnitude slower; B9 (DeepOPF) is
        the fastest but with lower margin (see Table~I).
  \item Parameter count scales modestly with grid size for
        JointOptimizer compared to MLP-based approaches.
\end{itemize}


% ============================================================
\section{Table VI -- Theorem 1 Validation}
\label{sec:table06}
% ============================================================

\input{tables/table06_theorem1.tex}

\medskip\noindent\textbf{What it shows.}
Empirical validation of Theorem~1:
$\rho(\tau) = |\lambda_{\min}(0)| - \sum_i K_i \cdot \tau_i / \tau_{\max,i}$.
For each delay level (50--500\,ms), the predicted margin $\rho_{\mathrm{pred}}$
is compared against the empirical stability rate across all 5 IEEE cases.

\noindent\textbf{Key points.}
\begin{itemize}
  \item The theoretical bound holds tightly across all grid sizes.
  \item Results use the auto-scaled $K_{\mathrm{init}}$, so the predicted
        margins differ per case according to $|\lambda_{\min}|$ and
        $n_{\mathrm{gen}}$.
  \item Stability degrades gracefully with increasing delay, consistent
        with the linear bound of Theorem~1.
\end{itemize}


% ============================================================
\section{Table VII -- N-1 Contingency Analysis}
\label{sec:table07}
% ============================================================

\input{tables/table07_n1_contingency.tex}

\medskip\noindent\textbf{What it shows.}
Stability rate and margin degradation under N-1 contingencies (single
line outages) for each IEEE case.  Reports average and worst-case
stability across all contingencies, plus the average margin after outage.

\noindent\textbf{Key points.}
\begin{itemize}
  \item Smaller grids (14--57) maintain 100\% stability under all
        single-line outages.
  \item IEEE 118 shows vulnerability under N-1 contingency, indicating
        that larger grids require contingency-aware training or
        additional margin reserves.
  \item This experiment is new to the journal extension and was not in
        the SmartGridComm submission.
\end{itemize}


% ============================================================
\section{Table VIII -- Delay Distribution Robustness}
\label{sec:table08}
% ============================================================

\input{tables/table08_delay_dist.tex}

\medskip\noindent\textbf{What it shows.}
Stability rate and margin under five different delay distributions:
lognormal, exponential, gamma, uniform, and Pareto.  Evaluated across
all 5 IEEE cases.

\noindent\textbf{Key points.}
\begin{itemize}
  \item Performance is robust across distribution types for cases
        14--57.
  \item Pareto delays (heavy-tailed) produce slightly higher margins on
        smaller grids because most samples have low delay.
  \item IEEE 118 is sensitive to delay distribution, with stability
        dropping significantly under lognormal and gamma tails.
  \item The Pareto distribution was added to \texttt{DelayConfig} via the
        \texttt{pareto\_alpha} field for this journal extension.
\end{itemize}


% ============================================================
\section{Table IX -- Convergence Analysis}
\label{sec:table09}
% ============================================================

\input{tables/table09_convergence.tex}

\medskip\noindent\textbf{What it shows.}
Training data efficiency: minimum number of scenarios needed for
$>$95\% stability, and stability margin at 100, 200, 300 training
scenarios, plus the best achieved margin.

\noindent\textbf{Key points.}
\begin{itemize}
  \item All cases reach $>$95\% stability with as few as 50 training
        scenarios.
  \item Margin continues to improve with more data, but with diminishing
        returns after 200 scenarios.
  \item IEEE 118 shows the largest gap between early and best margin,
        indicating that larger grids benefit more from additional
        training data.
\end{itemize}


% ============================================================
\section{Table X -- Model Compression}
\label{sec:table10}
% ============================================================

\input{tables/table10_compression.tex}

\medskip\noindent\textbf{What it shows.}
Effect of reducing the embedding dimension (and correspondingly the
hidden dimension and number of attention heads) on stability margin
and parameter count.

\noindent\textbf{Key points.}
\begin{itemize}
  \item Stability margin is remarkably stable across a wide range of
        model sizes (32K--1.8M parameters).
  \item A $15\times$ reduction in parameters (from $d=256$ to $d=32$)
        preserves the full stability margin.
  \item This suggests that JointOptimizer can be deployed on
        resource-constrained edge devices without performance loss.
\end{itemize}


% ============================================================
\section*{Summary of Changes from SmartGridComm 2026 Submission}
% ============================================================

\begin{enumerate}
  \item \textbf{5 IEEE cases} (14, 30, 39, 57, 118) instead of 3.
  \item \textbf{K-scaling fix}: $K_{\mathrm{init}} =
        \mathit{safety\_factor} \cdot |\lambda_{\min}| / n_{\mathrm{gen}}$
        instead of fixed $K=0.1$.
  \item \textbf{Theorem~2 renamed} to Observation~1 (information separation).
  \item \textbf{GPU training} (CUDA) instead of Apple M1 CPU.
  \item \textbf{New experiments} (Tables III, IV, VII, VIII, IX, X):
        stress test, transfer learning, N-1 contingency, delay
        distribution robustness, convergence analysis, model compression.
  \item \textbf{Additional baselines}: B8 (HeteroGNN) and B9 (DeepOPF)
        in the inference table.
  \item \textbf{Pad\'{e} analysis}: Pad\'{e}-approximant-based DDE
        simulation used for large-scale Theorem~1 validation (see
        companion figures, not tabulated).
  \item \textbf{Statistical rigor}: 5-seed runs with mean $\pm$ std
        reported in Tables I and II.
\end{enumerate}

\end{document}
